Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-01-16 Master Thesis/synopsis.tex
Clarify 4.2-pre

\begin{proof}
First we assume that $k/\mathbb{Q}$ is a Galois extension and $X$ has a $k$-rational point $P_0$. We can consider $k'$ as a subfield of $k$ via: 
\[\begin{tikzcd}
k'\ar[r] & k(P'_0)\ar[r, "\varphi_{P_0}"', "\cong"] & k(P_0) \ar[r, "\cong"] & k
\end{tikzcd}\]
For all $P\in X_0$ and $P'=\sigma(P)\in X'_0$ we also get an embedding:
\[ \begin{tikzcd}
k \ar[r] & k(P) \ar[r, "\varphi_P^{-1}"', "\cong"] & k(P')
\end{tikzcd} \]
We see any embedding $k\to \overline{k(P')}$ maps $k$ into $k(P')$ because $k/\mathbb{Q}$ is Galois. Thus we have for all $P'\in X'_0$:
\[ k(P')\otimes_{k'} k\cong \prod_{\operatorname{Hom}_{k'}(k, k(P'))} k(P') \]
in which case $X'\times_{k'}\operatorname{Spec} (k)\to X'$ splits completely. By Lemma \ref{4.1} we get $k\cong k'$, hence $[k:\mathbb{Q}]= [k':\mathbb{Q}]$.

Now we drop the assumption that $k/\mathbb{Q}$ is Galois. There exists a finite extension $k_1/k$ such that $k_1/\mathbb{Q}$ is Galois and $X(k_1)\neq\varnothing$. Let $K_1 = Kk_1$ be the constant field extension of $K$ with $k_1$ and let $k'_1$ be the constant field of $K'_1$. Applying the above argument to $K_1$ and $K'_1$ yields $[k_1:\mathbb{Q}] = [k'_1:\mathbb{Q}]$. Therefore:
\begin{align*}
[k:\mathbb{Q}] &= [k_1:\mathbb{Q}]\cdot [K_1:K]^{-1} = [k'_1:\mathbb{Q}]\cdot [K'_1:K']^{-1}\\
&= [k'_1:\mathbb{Q}]\cdot ([K'_1: K'k_1'] \cdot [K'k_1':K'])^{-1}\\
&= [k'_1:\mathbb{Q}]\cdot  ([K'_1: K'k_1'] \cdot [k'_1:k'])^{-1} \leq [k'_1:\mathbb{Q}]\cdot [k'_1:k']^{-1} = [k':\mathbb{Q}]
\end{align*}
By symmetry, we get $[k:\mathbb{Q}] = [k':\mathbb{Q}]$.
\end{proof}

>>>>

\begin{proof}
First we assume that $k/\mathbb{Q}$ is a Galois extension and $X$ has a $k$-rational point $P_0$. We can consider $k'$ as a subfield of $k$ via: 
\[\begin{tikzcd}
k'\ar[r] & k(P'_0)\ar[r, "\varphi_{P_0}"', "\cong"] & k(P_0) \ar[r, "\cong"] & k
\end{tikzcd}\]
For all $P\in X_0$ and $P'=\sigma(P)\in X'_0$ we also get an embedding:
\[ \begin{tikzcd}
k \ar[r] & k(P) \ar[r, "\varphi_P^{-1}"', "\cong"] & k(P')
\end{tikzcd} \]
Let us consider the morphism $f:X'\times_{k'} k\to X'$. We want to see that $f$ splits completely. For $P'\in X'_0$ we have as topological spaces:
\[ f^{-1}(P') = \operatorname{Spec} k(P')\times_{k'} k \]
Any embedding $k\to \overline{k(P')}$ maps $k$ into $k(P')$ because $k/\mathbb{Q}$ is Galois. Thus we have:
\[ k(P')\otimes_{k'} k\cong \prod_{\operatorname{Hom}_{k'}(k, k(P'))} k(P') \]
Now we can see that $f$ splits completely indeed. By Lemma \ref{4.1} we get $k\cong k'$, hence $[k:\mathbb{Q}]= [k':\mathbb{Q}]$.

Now we drop the assumption that $k/\mathbb{Q}$ is Galois. There exists a finite extension $k_1/k$ such that $k_1/\mathbb{Q}$ is Galois and $X(k_1)\neq\varnothing$. Let $K_1 = Kk_1$ be the constant field extension of $K$ with $k_1$ and let $k'_1$ be the constant field of $K'_1$. Applying the above argument to $K_1$ and $K'_1$ yields $[k_1:\mathbb{Q}] = [k'_1:\mathbb{Q}]$. Therefore:
\begin{align*}
[k:\mathbb{Q}] &= [k_1:\mathbb{Q}]\cdot [K_1:K]^{-1} \\
&= [k'_1:\mathbb{Q}]\cdot [K'_1:K']^{-1}\\
&= [k'_1:\mathbb{Q}]\cdot ([K'_1: K'k_1'] \cdot [K'k_1':K'])^{-1}\\
&= [k'_1:\mathbb{Q}]\cdot  ([K'_1: K'k_1'] \cdot [k'_1:k'])^{-1} \\
&\leq [k'_1:\mathbb{Q}]\cdot [k'_1:k']^{-1} = [k':\mathbb{Q}]
\end{align*}
By symmetry, we get $[k:\mathbb{Q}] = [k':\mathbb{Q}]$.
\end{proof}
